---
title: "Student test performance"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Drey Upsher-Obear"
output: html_document
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r, echo=FALSE}
library(tidyverse)
library(DescTools)
library(DataExplorer)
```

# Data 

Load data, check structure for variable types. 

```{r}
dat <- read.csv("StudentsPerformance.csv", sep = ",", stringsAsFactors = TRUE)

str(dat)
```


## Implied correlation

Using the wonderful DataExplorer package, lets check the distribution of our variables. 

```{r variable-dist, echo=FALSE}
plot_intro(dat)
```

A bit of a unicorn, no missing observations and no missing columns. Very little cleaning is necessary.

Now lets check for correlation between variables. 

```{r}
dat %>%
  plot_correlation(maxcat = 4)
```

Interesting, we have a pretty good idea of what to expect correlation wise with the variables.

## Question 1

Do students with mor educated parents have higher reading, writing, and mathematics scores? Are they more likely to study for the test given this?  

Step 1: organize data into a more readable form, then plot. Examining reading score first. I didn't want to reorganize the DF into a long form, so I will plot the 3 scores on separate plots. This will change eventually to a faceted graph. 

```{r parents-degree}
par.edu <- dat %>%
  group_by(parental.level.of.education) %>%
  summarise(avg_mat = mean(math.score),
            avg_writ = mean(writing.score),
            avg_read = mean(reading.score))

par.edu 

```

Plot for reading score. 

```{r reading-plot}
dat %>%
  ggplot(aes(parental.level.of.education, fill = parental.level.of.education)) + 
  geom_violin(aes(x = reorder(parental.level.of.education, -reading.score), y = reading.score)) + 
  coord_flip()
```

Plot for writing score.

```{r writing-plot}
dat %>%
  ggplot(aes(parental.level.of.education, fill = parental.level.of.education)) + 
  geom_violin(aes(x = reorder(parental.level.of.education, -writing.score), y = writing.score)) + 
  coord_flip()
```

Plot for mathematics score. 

```{r math-plot}
dat %>%
  ggplot(aes(parental.level.of.education, fill = parental.level.of.education)) + 
  geom_violin(aes(x = reorder(parental.level.of.education, -math.score), y = math.score)) + 
  coord_flip()
```

# Conclusion for question 1

From these plots, it would be plausible to accept our hypothesis. Students that have more educated parents do indeed score higher on reading, writing, and mathematics testing. 
